The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-09-14 Paper.Rmd
modified@Sept14

Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data. So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both incidence are rare (abundant) synchronously. If any one of them be environmental variable and the other is an observable (like abundance of a species) then it can also explain the causal relationship between them. Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets. In addition to a copula based \nameref{Model_selection}, we also carried out a \ref{SM-NPA} with three different statistics to verify our model selection based results.

@@

We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples. Details of each datasets used were given in \ref{SM-Data_detailed}.

@@

<!--### Nonparametric analysis of tail dependence \label{NPA}-->

In addition to the model selection approach described above, **nonparametric approaches** (for details see Appendix \ref{SM-NPA}) to quantify lower ($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics) and upper tail dependence ($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics) were also used to answer \hyperref[Intro]{\textbf{(Q1)}}.
Using two distinct methods, when results are consistent, provides greater confidence 
that results are robust.

<!--***DAN: There is either too little or too much here in this 
paragraph. The reader cannot get the main
idea, really, of what was done for these tests, from this paragraph alone - they will have to
look at the appendix. They need to be able to get the main idea. So you could either add detail,
or else take away detail and just say something like "We tested our statistics by testing whether 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed asymmetry of tail dependence 
known to be present in data simulated from various copulas (Appendix X)" and then stop there.
Then you would remove panels B-D from Fig. 1. At the moment I prefer this second option, but 
my preference is not strong, so if you prefer to keep the material and add detail that would be 
fine.-->
<!--**SG: Agreed with 2nd option and did things as necessary-->

<!--***DAN: I think the below material belongs in supp mat, because the reader can get the
main idea via the above text, and only some readers will be interested in the detailed 
algorithm. Please move it there, and then make the above links XYZ and ABC into live links
to the appropriate locations in the sup mat.

For bivariate datasets, surrogates were produced as follows.

- Given data $(x_i,y_i)$ for $i=1,\ldots,n$ and a 
one-parameter target copula family $\mathcal{B}(\theta)$ (here $\theta$ denotes the parameter), let $\tau$ be the
Kendall correlation of the $x_i$ with the $y_i$ and find the parameter value $\theta_\tau$ for which 
$\mathcal{B}(\theta_\tau)$ has Kendall correlation $\tau$ (this is possible for the one-parameter copula families of this 
study using the `iTau` function of the `VineCopula` package). 
- Generate data $(a_i,b_i)$, $i=1,\ldots,n$ from 
$\mathcal{B}(\theta_\tau)$. 
- Permute the ordered set $(x_1,\ldots,x_n)$ such that the permutation
$(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ has $\rank(x_{\sigma_x(i)})$ equal to $\rank(a_i)$ for all $i$, where 
$\rank(x_{\sigma_x(i)})$ is the rank of $x_{\sigma_x(i)}$ in the set $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$, and
$\rank(a_i)$ is the rank of $a_i$ in the set $(a_1,\ldots,a_n)$ (here $\sigma_x$ is a permutation of the 
indices $1,\ldots,n$). 
- Likewise permute $(y_1,\ldots,y_n)$
such that the permutation
$(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ has $\rank(y_{\sigma_y(i)})$ equal to $\rank(b_i)$ for all $i$ ($\sigma_y$ is 
another permutation of $1,\ldots,n$).
- The surrogate dataset is $(x_{\sigma_x(i)},y_{\sigma_y(i)})$ for $i=1,\ldots,n$.

Our code for this algorithm is [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R) in the 
[`BIVAN`](https://github.com/sghosh89/BIVAN) repository.

Because $(x_{\sigma_x(1)},\ldots,x_{\sigma_x(n)})$ 
is a permutation of $(x_1,\ldots,x_n)$ and $(y_{\sigma_y(1)},\ldots,y_{\sigma_y(n)})$ is a 
permutation 
of $(y_1,\ldots,y_n)$, the marginal distributions of the surrogate data are exactly the same as 
those of the original data. 
Because ranks of the surrogate data are matched to ranks of the $a_i$ and $b_i$, the
Kendall correlation of the surrogate dataset will be the same as that of the $a_i$ and $b_i$, and therefore 
will be close to $\tau$ (differences arising only through sampling variation). If instead of $\theta_\tau$
a parameter value $\theta_\rho$ is used such that the Spearman correlation of $\mathcal{B}(\theta_\rho)$ is the same
as that of the $x_i$ with the $y_i$, then Spearman correlations of surrogates will very nearly equal (except for 
sampling variation) the Spearman correlation of the empirical data.  
-->
<!--**SG: Done-->

@@

As described in Appendix \ref{SM-NPA}, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 Normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam}
contains a number which indicates how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.

@@

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in \ref{SM-NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.

@@

This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}. We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change? 
<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.--> 
<!--### The Moran effect and copula structure in populations-->
Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
variable between two locations, then there can be a non-normal copula structure for the
dependence of populations in the locations that are influenced by the environmental variables. This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations. We tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment and described our findings in Appendix \ref{SM-Cause_Moran}.
<!--
### Asymmetric sensitivity and copula structure in populations
Another model, distinct from the above, Dan and Shya to discuss-->
On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold. So, we considered another two-habitat patch model which was influenced by both local and global noises and found that a long time exposure to assymetric copula for global fluctuation shaped the population copula accordingly (for details, see Appendix \ref{SM-Cause_asym_sens}).

@@

This section is attributed to analyze of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data : we tested with four multivariate datasets and found results consistent with model selection and nonparametric approach (For details see Appendix \ref{SM-Consequence_skewness}). We then showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}). Finally, we studied **(iii)** the effects of non-Normal copula on extinction risk with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).

>>>>

Though a few examples have been found which shows copula approach to model dependencies in bioinformatics [@Kim2008], medical research [@Emura2016], a short period study of intersite dependence of rainfall data [@Serinaldi2008], a recent study on characterizing and assessing risks in drought management [@She2018] or monitoring changes in precipitation-extremes [@Goswami2018], the multivariate copula approach still, we believe, remains unexplored in ecology or related field and thus have potential scope to better capture the evolution of ecological time series data. So, in ecological perspective, when we find a dominantly lower (upper) tail dependence structure, it means both incidence are rare (abundant) synchronously. If any one of them be environmental variable and the other is an observable (like abundance of a species) then it can also explain the causal relationship between them. Here, we analyzed the dependence structure for several bivariate and multivariate ecological, community and biogeochemical datasets. In addition to a copula based \nameref{Model_selection}, we also carried out a \nameref{NPA} with three different statistics to verify our model selection based results.

@@

Considered datasets were ranged widely from population ecology, community ecology , evolutionery and biogeochemical sphere. 
We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples. Details of each dataset used were given in \ref{SM-Data_detailed}.

@@

### Nonparametric analysis of tail dependence \label{NPA}

In addition to the model selection approach described above, **nonparametric approaches** to quantify lower ($\cor_l$, $\Ps_l$ and $\Dsq_l$ statistics) and upper tail dependence ($\cor_u$, $\Ps_u$ and $\Dsq_u$ statistics) were also used to answer \hyperref[Intro]{\textbf{(Q1)}}. Using two distinct methods, when results are consistent, provides greater confidence 
that results are robust.

Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see \nameref{Model_selection} for defintions of $u_i$ and $v_i$)
are related to each other in the left and right tails, respectively, of their 
distributions [@nelsen2007_copula, @joe2014_dependence]. 
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
\end{center}
\end{figure}

The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
   (u_i-\mean(u)) (v_i-\mean(v)),
\end{equation}
where means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
of $h$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then defined
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
\end{equation}
As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
then small. 
Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
dependence in the portions of the distributions of $u$ and $v$ given by 
the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
$\Dsq_{l_b,u_b}$ indicate strong dependence.

For large datasets (i.e., datasets with large $n$),
we can consider $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we can consider 
several different bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.

The difference $\cor_l-\cor_u$ is expected to be positive if 
dependence in the left halves of the distributions is stronger 
than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
This holds, for instance, for the Clayton copula [@nelsen2007_copula].
The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
"lower" statistics occur in opposite order, is also expected to be 
positive under the same conditions. We use these three differences
to detect asymmetry of tail dependence. Likewise, for large datasets,
we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.

We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{SM-Test_npa_stat}).

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
but had Normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were Normal, but the data were otherwise 
statistically unchanged.
Details of how surrogates were produced are in Appendix \ref{SM-surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of question \hyperref[Intro]{\textbf{(Q1)}} 
from the \nameref{Intro} for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.

The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
as computed on the 
empirical data, to the distribution
of its values on many surrogate datasets 
provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
third part of question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro} for bivariate datasets.

For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
approach described above.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
Normal null hypothesis, and answering question \hyperref[Intro]{\textbf{(Q1)}} from the \nameref{Intro}.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].

@@

As described in \nameref{NPA}, each of our statistics, as computed for the 
ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (with $10$ number of bins), was compared to the distributions 
of values of the same statistic with the same ranges computed on 1000 Normal-copula 
surrogate datasets with similar Kendall or Spearman correlations, in different 
runs for Kendall and Spearman (Fig. \ref{SM-fig_soilCN_nonparam}). Each bin in each subplot of Fig. \ref{SM-fig_soilCN_nonparam}
contains a number which indicates how many surrogates out of 1000, are significantly different ($95\%$ confidence interval) from Normal-copula statistics.

@@

Mean values of our statistics ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$, 
$\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$), as well as mean Kendall and Spearman correlations, were 
computed across all pairs of sampling locations, and confidence intervals of these means were computed via a spatial
resampling scheme, as described in \nameref{NPA} (Table \ref{SM-tab_count_resamp}). Confidence intervals of
$\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, and $\Dsq_u$ always excluded zero and mean values were positive,
indicating tail dependence. Confidence intervals of $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$ also always 
exluded $0$ and were positive, indicating asymmetry of tail dependence (stronger lower tail dependence than upper), a feature that data with Normal copula structure do not have, providing further evidence answering the third part of question \hyperref[Intro]{\textbf{(Q1)}}.

@@

```{r read_RES_temp_aphid, echo=F}
RES_temp_aphid<-readRDS("./Results/fitting_results/AphidCopulaFit_selecloc_winter_neg_temp.RDS")
```

This section is dedicated to explore question \hyperref[Intro]{\textbf{(Q2)}} from \nameref{Intro}. We wanted to find how are copula structure of ecologically influential climate variables changing as part of climate change? 
<!--We now know the copula structure of environmental variables can induce (sometimes similar but not the same) copula structure in the populations, but we don’t really know how a change in the environment copula structure would make what kind of change in the population copula structure.--> 
<!--### The Moran effect and copula structure in populations-->
Firstly, Our hypothesis was if a non-normal copula structure describes the dependence of an environmental
variable between two locations, then there can be a non-normal copula structure for the
dependence of populations in the locations that are influenced by the environmental variables. This assumption was inspired by the the idea of Moran-effect [@Moran1953] which was originally used to explain synchronization of widely dispersed populations. We tested our hypothesis with a two habitat patch model (1st order AR method) in noisy environment (having non-Normal copula structure) and found that populations were shaped by similar copula structure as that of local noises ( for details see Appendix \ref{SM-Cause_Moran}). We also found the Moran effect on leaf-plum curling aphid's firstflight data induced by average winter temperature for the period of 56 years (1958-2013). We did pair wise model selection with negative average winter temperature (averaged over (n-1)$^{th}$ December to n$^{th}$ March temperature) from the selected 10 locations for which we did model selection with aphid-firstflight data. Negative winter temperature was considered for simulation to compare the aphid's firstflight results with temparature copula (aphid's firstflight date were negatively correlated with temperature or in other words positively correlated with negative temparature). Model selection results for negative average winter temperature in 10 $\times$ 10 matrices were represented in Table \ref{SM-tab_temp_aphid_bestcop} - Table \ref{SM-tab_temp_aphid_LTmUT}. In all cases `r min((RES_temp_aphid$gfc_numBS_success/RES_temp_aphid$gfc_numBS),na.rm=T)*100` $\%$ of bootstrap randomizations used for goodness of fit tests successfully lead to values of the statistic for comparison with its value on the real data. Table \ref{SM-tab_temp_aphid_LTmUT} showed `r sum(RES_temp_aphid$LTdep_AICw - RES_temp_aphid$UTdep_AICw <0, na.rm=T)` out of `r (dim(RES_temp_aphid$gfc_p_CvM)[1]^2-dim(RES_temp_aphid$gfc_p_CvM)[1])-2*RES_temp_aphid$num_indep_loc_pair` location pair showed upper tail dependence in temperature copula, which can be considered as the Moran effect results causing dominant upper tail dependence in the copulas with aphid's firstflight date (Table \ref{SM-tab_ff_LTmUT}).

On a second note, we intended to see if any specific (lower or upper) tail dependence emerges from a copula of two meta-populations, due to some global environmental factors (like temperature or snowfall) when it is beyond a certain expectation threshold. So, we considered another two-habitat patch model which was influenced by both local and global noises and found that a long time exposure to assymetric copula for global fluctuation shaped the population copula accordingly (for details, see Appendix \ref{SM-Cause_asym_sens}).

@@

This section is attributed to analyze of question \hyperref[Intro]{\textbf{(Q3)}} from \nameref{Intro}. We found that asymmetric tail dependence had **(i)** effects on skewness of spatially averaged time series data. Spatial mean of the time series data extracted that tail dependence from the multivariate data set, if any, and reflected in the averaged skewness. We tested null hypothesis $H_{0}$ which presumed that multivariate copula actually came from normal copula structure with no tail dependence and thus should have the same skewness as of normal copula (for details see Appendix \ref{SM-Consequence_skewness}). If $p$-value is less than $0.05$, then we reject the null hypothesis in support of significant tail dependence (and henceforth significantly different skewness). Figure \ref{fig_skewness_methane_maintext} showed asymmetric copula structure's effect on spatial skewness for biogeochemical methane-flux data.
The result, showing significantly different skewness of spatial mean from that of Normal surrogates, was consistent with model selection and nonparametric approach which supported lower tail dependence copula for methane-flux data.

We then showed **(ii)** asymmetric copula structure affected slope of spatial Taylor's Law differently than as expected from symmetric Normal copula (For details see Appendix \ref{SM-Consequence_TL}). Finally, we studied **(iii)** the effects of non-Normal copula on extinction risk with two metapopulation model, namely, Lewton-Cohen model and Ricker model (For details see Appendix \ref{SM-Consequence_ext_risk}).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/skewness_results/skewness_methane/skewness_methane.pdf}
\caption[Skewness-methane-flux data]{Skewness of spatially averaged methane-flux data (red line) was compared with respect to that of 10000 Normal surrogates preserving $(A)$ Kendall correlation, and $(B)$ Spearman correlation. p-value < 0.05 in both case showed significant difference in skewness (as expected for lower tail dependent methane flux data from model selection results) from Normal surrogates. \label{fig_skewness_methane_maintext}}
\end{center}
\end{figure}
